---
title: "R Data Structures and Operations Quiz"
output: webexercises::webexercises_default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(webexercises)
```
---

## Part 1: Vectors

### Question 1: Atomic Coercion
If you run `my_vec <- c(1, "A", TRUE)`, what will be the data type of the resulting vector?

```{r q1, results='asis'}
opts1 <- c(
  "A list containing a numeric, a character, and a logical value.",
  "An error, because vectors cannot contain mixed types.",
  answer = "A character vector: `\"1\"`, `\"A\"`, `\"TRUE\"`.",
  "A numeric vector with NAs introduced by coercion."
)
cat(longmcq(opts1))
```

`r hide("Show Explanation")`
Vectors must be "atomic" (all elements of the same type). R forces (coerces) them to the most flexible type. The hierarchy is Logical $\rightarrow$ Integer $\rightarrow$ Numeric $\rightarrow$ Character.
`r unhide()`

### Question 2: Vector Recycling
What happens when you add two vectors of different lengths, such as `c(1, 2, 3) + c(10, 20)`?

```{r q2, results='asis'}
opts2 <- c(
  "R throws a fatal error and stops execution.",
  "R adds the first two elements and drops the third: `11, 22`.",
  answer = "R recycles the shorter vector and issues a warning: `11, 22, 13`.",
  "R pads the shorter vector with zero: `11, 22, 3`."
)
cat(longmcq(opts2))
```

`r hide("Show Explanation")`
R uses "recycling" to repeat the shorter vector until it matches the length of the longer one. If the longer length isn't a multiple of the shorter one, it completes the operation but gives a warning.
`r unhide()`

### Question 3: Naming Vector Elements
Which function allows you to assign or retrieve labels for individual elements inside a vector?

```{r q3, results='asis'}
opts3 <- c(
  answer = "`names()`",
  "`labels()`",
  "`colnames()`",
  "`tags()`"
)
cat(longmcq(opts3))
```

`r hide("Show Explanation")`
The `names()` function is used for vectors and lists. (e.g., `names(my_vec) <- c("a", "b")`). `colnames()` is used for matrices and data frames.
`r unhide()`

---

## Part 2: Matrices & Arrays

### Question 4: Matrix Creation
By default, how does the `matrix()` function fill its data into rows and columns?

```{r q4, results='asis'}
opts4 <- c(
  "Row by row (horizontally).",
  answer = "Column by column (vertically).",
  "Randomly.",
  "Diagonally."
)
cat(longmcq(opts4))
```

`r hide("Show Explanation")`
By default, R fills matrices down the columns. To fill across the rows instead, you must explicitly set `byrow = TRUE`.
`r unhide()`

### Question 5: Matrix Properties
Which of the following is true about a matrix in R?

```{r q5, results='asis'}
opts5 <- c(
  answer = "It is a 2-dimensional structure that can only hold a single data type.",
  "It is a 2-dimensional structure that can hold mixed data types.",
  "It is a 1-dimensional structure used for linear algebra.",
  "It is simply a list of vectors."
)
cat(longmcq(opts5))
```

`r hide("Show Explanation")`
A matrix is essentially an atomic vector with a "dimensions" attribute (`dim`) attached to it. Because it is atomic, all data inside it must be of the exact same type.
`r unhide()`

### Question 6: Array Dimensions
What does the code `my_array <- array(1:24, dim = c(3, 4, 2))` create?

```{r q6, results='asis'}
opts6 <- c(
  "A 3D array with 2 rows, 4 columns, and 3 layers.",
  answer = "A 3D array with 3 rows, 4 columns, and 2 layers.",
  "A matrix with 3 rows and 8 columns.",
  "A list of 3 matrices, each 4x2."
)
cat(longmcq(opts6))
```

`r hide("Show Explanation")`
The `dim` argument for an array is ordered as `c(rows, columns, layers/tables)`. So `c(3, 4, 2)` means 3 rows, 4 columns, split across 2 separate "layers" or "slices".
`r unhide()`

### Question 7: Matrix Transposition
Which function is used to transpose a matrix (swap its rows and columns)?

```{r q7, results='asis'}
opts7 <- c(
  "`transpose()`",
  "`swap()`",
  "`flip()`",
  answer = "`t()`"
)
cat(longmcq(opts7))
```

`r hide("Show Explanation")`
The `t()` function returns the transpose of a matrix or a data frame.
`r unhide()`

---

## Part 3: Factors

### Question 8: Purpose of Factors
What is the primary purpose of the `factor` data structure in R?

```{r q8, results='asis'}
opts8 <- c(
  "To handle complex mathematical factorials.",
  "To store highly compressed textual data.",
  answer = "To store categorical data and assign distinct 'levels' to groups.",
  "To convert matrices into 1D vectors."
)
cat(longmcq(opts8))
```

`r hide("Show Explanation")`
Factors are used for categorical variables (like "Small", "Medium", "Large", or "Male", "Female"). R stores them as integers under the hood, mapped to character labels called "levels".
`r unhide()`

### Question 9: Ordered Factors
How do you tell R that a factor has a specific, meaningful order (e.g., Low < Medium < High)?

```{r q9, results='asis'}
opts9 <- c(
  "Use `sort(factor_data)`.",
  answer = "Use the `ordered = TRUE` argument inside the `factor()` function.",
  "Use the `levels()` function to automatically sort them.",
  "Factors are always ordered alphabetically by default; you cannot change it."
)
cat(longmcq(opts9))
```

`r hide("Show Explanation")`
Setting `ordered = TRUE` makes it an ordinal categorical variable, which is highly important for statistical modeling and plotting.
`r unhide()`

### Question 10: Converting Factors to Numeric
If `f <- factor(c("10", "20", "30"))`, what is the safest way to convert it back to actual numeric values?

```{r q10, results='asis'}
opts10 <- c(
  "`as.numeric(f)`",
  answer = "`as.numeric(as.character(f))` or `as.numeric(levels(f))[f]`",
  "`numeric(f)`",
  "`unfactor(f)`"
)
cat(longmcq(opts10))
```

`r hide("Show Explanation")`
If you just use `as.numeric(f)`, R returns the underlying integer codes (1, 2, 3) instead of the text values (10, 20, 30). You must convert it to a character first.
`r unhide()`

---

## Part 4: Lists & Data Frames

### Question 11: The Nature of Lists
Which statement is TRUE regarding lists in R?

```{r q11, results='asis'}
opts11 <- c(
  "Lists can only contain one data type.",
  "Lists are exactly the same as data frames.",
  answer = "Lists can contain elements of different types, lengths, and even other lists.",
  "Lists are always two-dimensional."
)
cat(longmcq(opts11))
```

`r hide("Show Explanation")`
Lists are the most flexible data structure in R. They are 1-dimensional and heterogeneous, meaning a single list can hold a vector, a matrix, a function, and another list all at once.
`r unhide()`

### Question 12: Single vs Double Brackets
If `my_list` contains a vector named `scores` at the first position, what is the difference between `my_list[1]` and `my_list[[1]]`?

```{r q12, results='asis'}
opts12 <- c(
  "They are completely identical.",
  answer = "`[1]` returns a list of length 1, while `[[1]]` extracts the actual vector of scores out of the list.",
  "`[1]` extracts the vector, while `[[1]]` returns a list.",
  "`[[1]]` only works for data frames, not lists."
)
cat(longmcq(opts12))
```

`r hide("Show Explanation")`
Think of a list as a train with cargo cars. Single brackets `[]` give you a train car (still a list). Double brackets `[[]]` open the door and give you the cargo inside.
`r unhide()`

### Question 13: Data Frame Constraints
Which of the following is a strict requirement for creating a valid Data Frame?

```{r q13, results='asis'}
opts13 <- c(
  "All columns must be of the same data type.",
  answer = "All columns must have the exact same number of rows (length).",
  "It cannot contain factor variables.",
  "Row names must be strictly numeric."
)
cat(longmcq(opts13))
```

`r hide("Show Explanation")`
A data frame is technically a list of vectors. For it to form a neat rectangle (table), every vector (column) must be exactly the same length.
`r unhide()`

### Question 14: Adding a Column
Which syntax is commonly used to add a brand new column named `Total` to an existing data frame `df`?

```{r q14, results='asis'}
opts14 <- c(
  answer = "`df$Total <- c(...)`",
  "`df[Total] <- c(...)`",
  "`add_col(df, Total)`",
  "`df %>% Total(...)`"
)
cat(longmcq(opts14))
```

`r hide("Show Explanation")`
The `$` operator is the easiest way to extract or create columns. Assigning values to a `$` name that doesn't exist yet will automatically create it.
`r unhide()`

### Question 15: Examining Structure
Which function gives you a compact summary of a data frame's internal structure, showing column names, types, and the first few observations?

```{r q15, results='asis'}
opts15 <- c(
  "`summary()`",
  "`head()`",
  answer = "`str()`",
  "`class()`"
)
cat(longmcq(opts15))
```

`r hide("Show Explanation")`
`str()` stands for "structure". It is one of the most useful diagnostic functions in R for checking the types of data stored in your data frame.
`r unhide()`

---

## Part 5: Indexing & Conditional Selection

### Question 16: Negative Indexing
What does `my_vector[-3]` do?

```{r q16, results='asis'}
opts16 <- c(
  "It returns the 3rd element from the end of the vector.",
  answer = "It returns the entire vector except for the 3rd element.",
  "It reverses the vector.",
  "It subtracts 3 from every element in the vector."
)
cat(longmcq(opts16))
```

`r hide("Show Explanation")`
In R, placing a minus sign before an index tells R to drop or exclude that specific index and return everything else.
`r unhide()`

### Question 17: Matrix Indexing
How do you extract the entire 2nd column of a matrix named `mat`?

```{r q17, results='asis'}
opts17 <- c(
  "`mat[2, ]`",
  answer = "`mat[, 2]`",
  "`mat[2]`",
  "`mat[[2]]`"
)
cat(longmcq(opts17))
```

`r hide("Show Explanation")`
Matrix indexing uses `[rows, columns]`. Leaving the row slot blank (before the comma) tells R to return "all rows" for column 2.
`r unhide()`

### Question 18: Out of Bounds
If `x <- c(10, 20, 30)`, what does R return if you ask for `x[5]`?

```{r q18, results='asis'}
opts18 <- c(
  "An 'Index out of bounds' error.",
  "0",
  "NULL",
  answer = "NA"
)
cat(longmcq(opts18))
```

`r hide("Show Explanation")`
Unlike Python which crashes with an index error, R silently returns `NA` (Not Available) when you try to access an index that doesn't exist.
`r unhide()`

### Question 19: Logical Vectors
If `v <- c(1, 5, 10)`, what is the result of the expression `v > 4`?

```{r q19, results='asis'}
opts19 <- c(
  "It filters the vector and returns `5, 10`.",
  answer = "It returns a logical vector: `FALSE, TRUE, TRUE`.",
  "It returns the indices: `2, 3`.",
  "It returns `TRUE`."
)
cat(longmcq(opts19))
```

`r hide("Show Explanation")`
A logical condition applied to a vector is evaluated element-by-element, returning a boolean vector of the same length. You can use this logical vector to subset the data (e.g., `v[v > 4]`).
`r unhide()`

### Question 20: The `which()` Function
What does the `which()` function return when applied to a logical condition?

```{r q20, results='asis'}
opts20 <- c(
  "The actual data values that meet the condition.",
  "A boolean vector.",
  answer = "The numeric index positions of the `TRUE` values.",
  "A subset data frame."
)
cat(longmcq(opts20))
```

`r hide("Show Explanation")`
`which(v > 4)` would return `c(2, 3)` because the 2nd and 3rd elements satisfy the condition.
`r unhide()`

### Question 21: AND vs OR
Which operator is used to evaluate if *either* condition A OR condition B is true element-wise in a vector?

```{r q21, results='asis'}
opts21 <- c(
  "`&`",
  "`&&`",
  answer = "`|`",
  "`||`"
)
cat(longmcq(opts21))
```

`r hide("Show Explanation")`
The single vertical bar `|` is the vectorized OR operator. (The single `&` is the vectorized AND operator). The double `||` and `&&` only evaluate the very first element, usually used in `if()` statements.
`r unhide()`

### Question 22: Using `subset()`
What is the main advantage of using the `subset(df, condition)` function over standard bracket `df[condition, ]` subsetting?

```{r q22, results='asis'}
opts22 <- c(
  "It is much faster for large datasets.",
  answer = "It saves typing the data frame name repeatedly and automatically drops `NA` values in the condition.",
  "It can be used on lists, whereas brackets cannot.",
  "It modifies the data frame in place permanently."
)
cat(longmcq(opts22))
```

`r hide("Show Explanation")`
Instead of typing `df[df$Age > 20 & !is.na(df$Age), ]`, you can just type `subset(df, Age > 20)`. It is cleaner and handles missing values nicely.
`r unhide()`

### Question 23: The `%in%` Operator
What is the `%in%` operator used for?

```{r q23, results='asis'}
opts23 <- c(
  "To check if a package is installed.",
  "To concatenate two strings together.",
  answer = "To check if the elements of one vector match any of the elements in another vector.",
  "To pipe data from one function to another."
)
cat(longmcq(opts23))
```

`r hide("Show Explanation")`
`x %in% y` returns a logical vector the same length as `x`, marking `TRUE` if that element is found anywhere inside `y`.
`r unhide()`

---

## Part 6: Missing Values (NA)

### Question 24: Testing for NA
Why does `x == NA` fail to properly identify missing values?

```{r q24, results='asis'}
opts24 <- c(
  answer = "Because NA represents an unknown value, so checking if an unknown equals an unknown results in NA, not TRUE/FALSE.",
  "Because NA is a character string and requires quotes (`\"NA\"`).",
  "Because R uses `NULL` instead of `NA`.",
  "Because the syntax should be `x = NA`."
)
cat(longmcq(opts24))
```

`r hide("Show Explanation")`
You can never use equality operators with `NA`. You must use the dedicated function `is.na(x)` to test for missing values.
`r unhide()`

### Question 25: Calculating with NAs
If `x <- c(10, 20, NA, 40)`, what does `mean(x)` return?

```{r q25, results='asis'}
opts25 <- c(
  "23.33",
  "17.5",
  "An error.",
  answer = "NA"
)
cat(longmcq(opts25))
```

`r hide("Show Explanation")`
Most mathematical functions in R will return `NA` if even one `NA` is present in the data. You must add the argument `na.rm = TRUE` to strip them out before calculation.
`r unhide()`

### Question 26: Complete Cases
Which function evaluates a data frame and returns a logical vector marking `TRUE` only for rows that have absolutely zero missing values?

```{r q26, results='asis'}
opts26 <- c(
  answer = "`complete.cases()`",
  "`na.omit()`",
  "`is.complete()`",
  "`drop_na()`"
)
cat(longmcq(opts26))
```

`r hide("Show Explanation")`
`complete.cases(df)` returns a boolean vector you can use to subset the data frame (e.g., `df[complete.cases(df), ]`). Alternatively, `na.omit(df)` does the subsetting automatically.
`r unhide()`

---

## Part 7: Grouped Data 

### Question 27: The `tapply()` Function
What is `tapply()` used for?

```{r q27, results='asis'}
opts27 <- c(
  "Transposing a data frame.",
  answer = "Applying a function to subsets of a vector, split by a categorical grouping variable.",
  "Testing for missing values across rows.",
  "Tabulating frequencies of text strings."
)
cat(longmcq(opts27))
```

`r hide("Show Explanation")`
`tapply(X, INDEX, FUN)` takes a vector `X`, splits it into groups based on the factor `INDEX`, and applies `FUN` (like `mean`) to each group.
`r unhide()`

### Question 28: Splitting Data
Which function breaks a data frame into a *list* of smaller data frames based on a grouping variable?

```{r q28, results='asis'}
opts28 <- c(
  answer = "`split()`",
  "`group_by()`",
  "`divide()`",
  "`chunk()`"
)
cat(longmcq(opts28))
```

`r hide("Show Explanation")`
`split(df, df$category)` returns a list where each element is a subset of the original data frame for a specific category level.
`r unhide()`

### Question 29: The `aggregate()` Function
How does the `aggregate()` function handle groupings compared to `tapply()`?

```{r q29, results='asis'}
opts29 <- c(
  "It can only calculate sums, not means or standard deviations.",
  answer = "It natively accepts a formula interface (e.g., `y ~ group`) and returns a neat data frame instead of an array.",
  "It is strictly used for time-series data.",
  "It modifies the original dataset by adding summary rows at the bottom."
)
cat(longmcq(opts29))
```

`r hide("Show Explanation")`
`aggregate(value ~ category, data = df, FUN = mean)` is highly preferred in base R because its formula syntax is easy to read, and it outputs a clean data frame.
`r unhide()`

### Question 30: By Function
The `by()` function is an object-oriented wrapper for `tapply()`. What is it most commonly used to do?

```{r q30, results='asis'}
opts30 <- c(
  "Sort a data frame by multiple columns.",
  answer = "Apply a function to entire data frames split by row factors.",
  "Merge two datasets by a common ID column.",
  "Generate sequential numbers by a specific step size."
)
cat(longmcq(opts30))
```

`r hide("Show Explanation")`
While `tapply()` works on individual vectors, `by(data, INDICES, FUN)` splits an entire data frame into subsets and allows you to apply a function (like a linear model `lm()`) to each subset.
`r unhide()`

---

### Question 31: Vector Indexing Out-of-Bounds
What value does R return if you attempt to access an index position in a vector that does not exist (e.g., `x[10]` on a vector of length 5)?

```{r q31, echo = FALSE}
opts31 <- c(
  answer = "`NA`<br>",
  "`NULL`<br>",
  "An error message<br>",
  "`0`<br>"
)
```

`r longmcq(opts31)`

---

### Question 32: Checking Data Types
Which function should you use to check the basic data type or internal structure of an R object (e.g., whether a variable is `numeric`, `character`, or `logical`)?

```{r q32, echo = FALSE}
opts32 <- c(
  answer = "`class()`<br>",
  "`type()`<br>",
  "`datatype()`<br>",
  "`kind()`<br>"
)
```

`r longmcq(opts32)`

---

---
### Question 31: Combining Vectors into a Matrix
Which function binds a set of equal-length vectors together as columns to form a matrix?

```{r 31, echo = FALSE}
opts31<- c(
  answer = "`cbind()`<br>",
  "`rbind()`<br>",
  "`bind_cols()`<br>",
  "`concat()`<br>"
)
```

`r longmcq(opts31)`

---

### Question 32: Removing Missing Values
When calculating the mean of a vector that contains `NA` values, which argument must be set to `TRUE` to ignore the missing values?

```{r 32, echo = FALSE}
opts32 <- c(
  answer = "`na.rm`<br>",
  "`remove.na`<br>",
  "`ignore.na`<br>",
  "`na.omit`<br>"
)
```

`r longmcq(opts32)`

---

### Question 33: List Element Extraction
If you have a list named `my_list`, which syntax correctly extracts the underlying object stored in the second position rather than returning a sub-list?

```{r 33, echo = FALSE}
opts33 <- c(
  answer = "`my_list[[2]]`<br>",
  "`my_list[2]`<br>",
  "`my_list(2)`<br>",
  "`my_list$2`<br>"
)
```

`r longmcq(opts33)`

---

### Question 34: Identifying Object Names
Which function displays the names of all objects (variables, vectors, functions) currently loaded in your active R workspace?

```{r 34, echo = FALSE}
opts34 <- c(
  answer = "`ls()`<br>",
  "`objects.list()`<br>",
  "`dir()`<br>",
  "`show()`<br>"
)
```

`r longmcq(opts34)`

---

### Question 35: Subsetting Vectors with Logic
Given the vector `x <- c(10, 25, 5, 30)`, which command returns only the values that are strictly greater than 15?

```{r 35, echo = FALSE}
opts35 <- c(
  answer = "`x[x > 15]`<br>",
  "`x(x > 15)`<br>",
  "`subset(x > 15)`<br>",
  "`x[[x > 15]]`<br>"
)
```

`r longmcq(opts35)`

---

### Question 36: Logical Operators in Subsetting
Which logical operator should you use inside vector brackets `[]` to select elements that satisfy both conditions simultaneously (e.g., values greater than 5 and less than 20)?

```{r 36, echo = FALSE}
opts36 <- c(
  answer = "`&`<br>",
  "`&&`<br>",
  "`AND`<br>",
  "`|`<br>"
)
```

`r longmcq(opts36)`
---

### Question 37: Filtering Data Frame Rows by Condition
Given a data frame df with a numeric column named age, which command filters the data frame to keep only rows where `age` is 18 or older?

```{r 37, echo = FALSE}
opts37 <- c(
  answer = "`df[df$age >= 18, ]`<br>",
  "`df[df$age >= 18]`<br>",
  "`df[ , df$age >= 18]`<br>",
  "`df[[df$age >= 18]]`<br>"
)
```

`r longmcq(opts37)`
---

### Question 38: Using the `subset()` Function
What is the main advantage of using `subset(df, age >= 18)` over bracket indexing `df[df$age >= 18, ]` when working with data frames?

```{r 38, echo = FALSE}
opts38 <- c(
  answer = "It automatically excludes rows where the condition evaluates to `NA`<br>",
  "It is faster and uses less computer memory<br>",
  "It works on matrices, whereas bracket indexing does not<br>",
  "It permanently modifies the original data frame<br>"
)
```

`r longmcq(opts38)`

---

### Question 39: Negative Indexing in Vectors
What does the command `temperature[-1]` do when applied to a numeric vector named `temperature`?

```{r 39, echo = FALSE}
opts39 <- c(
  answer = "It returns all elements of the vector except the first one<br>",
  "It subtracts 1 from the first element of the vector<br>",
  "It returns only the last element of the vector<br>",
  "It replaces the first element with `NA`<br>"
)
```

`r longmcq(opts39)`

---

### Question 40:Conditional Subsetting on List Named Elements
If a list `my_list` contains named elements `c("A", "B", "C")`, which syntax safely extracts the element named "B" by name?

```{r 40, echo = FALSE}
opts40 <- c(
  answer = "`my_list[[\"B\"]]`<br>",
  "`my_list[\"B\"]`<br>",
  "`my_list[B]`<br>",
  "`my_list(B)`<br>"
)
```

`r longmcq(opts40)`

---