---
title: "Probability"
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(webexercises)
```

# Welcome to the Probability Quiz!

Test your knowledge of random distributions, probability functions, and simulations in Base R. Select the correct answer for each question below.

---

### Question 1: Reproducibility in Probability
Which function is used to ensure that a sequence of random numbers generated in R can be exactly reproduced by another user?

```{r q1, results='asis'}
opts1 <- c(
  "`random.seed()`<br>",
  answer = "`set.seed()`<br>",
  "`reproduce()`<br>",
  "`fix.seed()`<br>"
)
```

`r webexercises::longmcq(opts1)`

`r hide("Show Explanation")`
Using `set.seed(123)` (or any integer) forces R's random number generator to start at the exact same sequence, making simulations fully reproducible.
`r unhide()`

---

### Question 2: The Prefix Convention
In R, probability distributions use a standard set of prefixes (`d`, `p`, `q`, `r`). Which prefix is used to calculate the **Cumulative Distribution Function (CDF)**?

```{r q2, results='asis'}
opts2 <- c(
  "`d` (e.g., `dnorm`)<br>",
  answer = "`p` (e.g., `pnorm`)<br>",
  "`q` (e.g., `qnorm`)<br>",
  "`r` (e.g., `rnorm`)<br>"
)
cat(longmcq(opts2))
```

`r hide("Show Explanation")`
The `p` prefix stands for "probability" and calculates the CDF (the probability that a random variable is less than or equal to a value). `d` is for density, `q` is for quantile, and `r` is for random generation.
`r unhide()`

---

### Question 3: Simulating a Die Roll
How would you simulate rolling a fair six-sided die 10 times in R?

```{r q3, results='asis'}
opts3 <- c(
  answer = "`sample(1:6, 10, replace = TRUE)`<br>",
  "`sample(1:6, 10, replace = FALSE)`<br>",
  "`runif(10, 1, 6)`<br>",
  "`rbinom(10, 6, 0.5)`<br>"
)
cat(longmcq(opts3))
```

`r hide("Show Explanation")`
Because a die can land on the same number multiple times, you must sample from the numbers 1 to 6 **with replacement** (`replace = TRUE`).
`r unhide()`

---

### Question 4: Normal Distribution Density
Which function returns the height of the probability density function (PDF) for a normal distribution at a specific value `x`?

```{r q4, results='asis'}
opts4 <- c(
  answer = "`dnorm(x)`<br>",
  "`pnorm(x)`<br>",
  "`qnorm(x)`<br>",
  "`rnorm(x)`<br>"
)
cat(longmcq(opts4))
```

`r hide("Show Explanation")`
The `d` prefix stands for "density", which evaluates the PDF curve at the given coordinate.
`r unhide()`

---

### Question 5: Generating Normal Random Variables
What does the code `rnorm(5, mean = 10, sd = 2)` do?

```{r q5, results='asis'}
opts5 <- c(
  "Generates 5 random numbers from a uniform distribution between 10 and 2.<br>",
  "Generates 5 random numbers from a normal distribution with a mean of 10 and a variance of 2.<br>",
  answer = "Generates 5 random numbers from a normal distribution with a mean of 10 and a standard deviation of 2.<br>",
  "Calculates the exact probability of getting the number 5 in a normal distribution.<br>"
)
cat(longmcq(opts5))
```

`r hide("Show Explanation")`
The `rnorm()` function takes the number of observations (`n`), the `mean`, and the standard deviation (`sd`) as its primary arguments.
`r unhide()`

---

### Question 6: Uniform Distribution
Which command generates a single random number uniformly distributed between 0 and 1?

```{r q6, results='asis'}
opts6 <- c(
  answer = "`runif(1, min = 0, max = 1)`<br>",
  "`rnorm(1, 0, 1)`<br>",
  "`sample(0:1, 1)`<br>",
  "`rbinom(1, 1, 0.5)`<br>"
)
cat(longmcq(opts6))
```

`r hide("Show Explanation")`
`runif()` stands for "random uniform." It generates a continuous decimal value equally likely to fall anywhere between the `min` and `max` parameters.
`r unhide()`

---

### Question 7: Exact Binomial Probability
How do you calculate the exact probability of getting exactly 3 heads in 10 flips of a fair coin?

```{r q7, results='asis'}
opts7 <- c(
  "`pbinom(3, size = 10, prob = 0.5)`<br>",
  answer = "`dbinom(3, size = 10, prob = 0.5)`<br>",
  "`qbinom(3, size = 10, prob = 0.5)`<br>",
  "`rbinom(3, size = 10, prob = 0.5)`<br>"
)
cat(longmcq(opts7))
```

`r hide("Show Explanation")`
For discrete distributions like the binomial, the density function (`d` prefix) calculates the exact probability of a specific outcome $P(X = x)$.
`r unhide()`

---

### Question 8: Cumulative Binomial Probability
Which function calculates the probability of getting **3 or fewer** heads in 10 flips of a fair coin?

```{r q8, results='asis'}
opts8 <- c(
  answer = "`pbinom(3, size = 10, prob = 0.5)`<br>",
  "`dbinom(3, size = 10, prob = 0.5)`<br>",
  "`sum(rbinom(10, 1, 0.5) <= 3)`<br>",
  "`qbinom(3, size = 10, prob = 0.5)`<br>"
)
cat(longmcq(opts8))
```

`r hide("Show Explanation")`
The `p` prefix (CDF) sums up the probabilities of getting 0, 1, 2, and 3 heads $P(X \le 3)$.
`r unhide()`

---

### Question 9: Poisson Distribution
If a call center receives an average of 5 calls per hour, which function finds the probability of receiving exactly 7 calls in a specific hour?

```{r q9, results='asis'}
opts9 <- c(
  answer = "`dpois(7, lambda = 5)`<br>",
  "`ppois(7, lambda = 5)`<br>",
  "`dpois(5, lambda = 7)`<br>",
  "`rpois(7, lambda = 5)`<br>"
)
cat(longmcq(opts9))
```

`r hide("Show Explanation")`
The `dpois()` function evaluates the probability mass for the Poisson distribution. The `lambda` argument represents the known average rate.
`r unhide()`

---

### Question 10: Combinatorics
Which R function calculates the number of combinations (e.g., "10 choose 3")?

```{r q10, results='asis'}
opts10 <- c(
  "`comb(10, 3)`<br>",
  answer = "`choose(10, 3)`<br>",
  "`factorial(10) / factorial(3)`<br>",
  "`combinations(10, 3)`<br>"
)
cat(longmcq(opts10))
```

`r hide("Show Explanation")`
The `choose(n, k)` function is a built-in base R function that computes the binomial coefficient, avoiding manual factorial arithmetic.
`r unhide()`

---

### Question 11: Weighted Sampling
How do you draw a single item from the vector `c('A', 'B')`, where 'A' has a 70% probability of being chosen and 'B' has a 30% probability?

```{r q11, results='asis'}
opts11 <- c(
  answer = "`sample(c('A', 'B'), 1, prob = c(0.7, 0.3))`<br>",
  "`sample(c('A', 'B'), 1, weights = c(70, 30))`<br>",
  "`rbinom(1, size = 2, prob = 0.7)`<br>",
  "`choose(c('A', 'B'), 1, prob = c(0.7, 0.3))`<br>"
)
cat(longmcq(opts11))
```

`r hide("Show Explanation")`
The `sample()` function accepts a `prob` argument where you can pass a vector of custom probabilities corresponding to each item you are sampling.
`r unhide()`

---

### Question 12: Finding Quantiles
Which function finds the Z-score corresponding to the 95th percentile of the standard normal distribution?

```{r q12, results='asis'}
opts12 <- c(
  answer = "`qnorm(0.95)`<br>",
  "`pnorm(0.95)`<br>",
  "`dnorm(0.95)`<br>",
  "`rnorm(0.95)`<br>"
)
cat(longmcq(opts12))
```

`r hide("Show Explanation")`
The `q` prefix stands for "quantile." It is the inverse of the CDF. If you give it a probability (0.95), it returns the threshold value (Z-score) below which that percentage of the data lies.
`r unhide()`

---

### Question 13: Upper Tail Probabilities
By default, `pnorm(1.96)` calculates the lower tail $P(X \le 1.96)$. How do you calculate the upper tail probability $P(X > 1.96)$ using the same function?

```{r q13, results='asis'}
opts13 <- c(
  "`pnorm(1.96, upper = TRUE)`<br>",
  answer = "`pnorm(1.96, lower.tail = FALSE)`<br>",
  "`1 + pnorm(1.96)`<br>",
  "`pnorm(1.96, tail = 'right')`<br>"
)
cat(longmcq(opts13))
```

`r hide("Show Explanation")`
Setting the `lower.tail` argument to `FALSE` is the standard way to calculate upper tail probabilities in R, which is mathematically equivalent to `1 - pnorm(1.96)`.
`r unhide()`

---

### Question 14: The Exponential Distribution
What does `rexp(10, rate = 0.2)` generate?

```{r q14, results='asis'}
opts14 <- c(
  answer = "10 random values from an exponential distribution with lambda = 0.2.<br>",
  "The probability density of an exponential distribution evaluated at 10.<br>",
  "10 random values from a Poisson distribution.<br>",
  "The cumulative probability up to the value 10.<br>"
)
cat(longmcq(opts14))
```

`r hide("Show Explanation")`
The `r` prefix generates random sequences, and `exp` references the exponential distribution, which models the time between events using the `rate` parameter.
`r unhide()`

---

### Question 15: Factorials in Math
To calculate $5!$ (5 factorial), which R code should you use?

```{r q15, results='asis'}
opts15 <- c(
  "`fact(5)`<br>",
  answer = "`factorial(5)`<br>",
  "`prod(1, 5)`<br>",
  "`5!`<br>"
)
cat(longmcq(opts15))
```

`r hide("Show Explanation")`
R has a dedicated `factorial()` function. Using the literal exclamation mark (`5!`) will result in a syntax error because `!` is used as the logical "NOT" operator in R.
`r unhide()`

---

### Question 16: Empirical Probabilities
If `x <- runif(100)`, what does `mean(x < 0.2)` approximately evaluate to?

```{r q16, results='asis'}
opts16 <- c(
  "The exact number 0.2.<br>",
  "A logical vector of TRUE and FALSE.<br>",
  answer = "The empirical proportion of generated values less than 0.2 (approximately 0.2).<br>",
  "An error because `mean()` cannot take a logical condition.<br>"
)
cat(longmcq(opts16))
```

`r hide("Show Explanation")`
When you take the mean of a logical statement (like `x < 0.2`), R treats `TRUE` as 1 and `FALSE` as 0. Therefore, the average equals the proportion of times the condition was met.
`r unhide()`

---

### Question 17: Default Sampling Behavior
What happens if you run `sample(1:50, 5)`?

```{r q17, results='asis'}
opts17 <- c(
  "It draws 5 numbers from 1 to 50 with replacement.<br>",
  answer = "It draws 5 numbers from 1 to 50 without replacement.<br>",
  "It samples numbers from 1 to 5 exactly 50 times.<br>",
  "It returns an error because `replace` is missing.<br>"
)
cat(longmcq(opts17))
```

`r hide("Show Explanation")`
By default, the `sample()` function always operates **without replacement** (`replace = FALSE`). You must explicitly change this if you want duplicate draws.
`r unhide()`

---

### Question 18: Student's t-Distribution
Which function finds the cumulative probability $P(T \le 2.1)$ for a t-distribution with 15 degrees of freedom?

```{r q18, results='asis'}
opts18 <- c(
  answer = "`pt(2.1, df = 15)`<br>",
  "`dt(2.1, df = 15)`<br>",
  "`qt(2.1, df = 15)`<br>",
  "`rt(2.1, df = 15)`<br>"
)
cat(longmcq(opts18))
```

`r hide("Show Explanation")`
The `p` prefix is used for the cumulative probability, and `t` represents the t-distribution. The degrees of freedom are set using the `df` argument.
`r unhide()`

---

### Question 19: Chi-Squared Simulation
How do you generate 20 random observations from a Chi-squared distribution with 4 degrees of freedom?

```{r q19, results='asis'}
opts19 <- c(
  answer = "`rchisq(20, df = 4)`<br>",
  "`rchi(20, df = 4)`<br>",
  "`qchisq(20, df = 4)`<br>",
  "`pchisq(20, df = 4)`<br>"
)
cat(longmcq(opts19))
```

`r hide("Show Explanation")`
The function name for the Chi-squared distribution in R is `chisq`. Adding the `r` prefix allows you to generate random values from it.
`r unhide()`

---

### Question 20: Calculating Empirical Probability
You simulate 1000 coin flips with `flips <- sample(c('H', 'T'), 1000, replace = TRUE)`. Which of the following correctly calculates the empirical probability of getting heads ('H')?

```{r q20, results='asis'}
opts20 <- c(
  "`sum(flips == 'H') / 1000`<br>",
  "`mean(flips == 'H')`<br>",
  answer = "Both `sum(flips == 'H') / 1000` and `mean(flips == 'H')` will work.<br>",
  "`prop(flips, 'H')`<br>"
)
cat(longmcq(opts20))
```

`r hide("Show Explanation")`
Summing the true statements and dividing by the total `sum(x) / length(x)` is mathematically identical to just calculating the average of the true statements `mean(x)`. Both are completely valid ways to calculate probability in simulations.
`r unhide()`