---
title: "Introduction to R and RStudio"
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(webexercises)
```


```{r, echo = FALSE}
# Uncomment to change widget colours:
#style_widgets(incorrect = "goldenrod", correct = "purple")
```


## Longer MCQs (`longmcq()`)

When your answers are very long, sometimes a drop-down select box gets formatted oddly. You can use `longmcq()` to deal with this. Since the answers are long, It's probably best to set up the options inside an R chunk with `echo=FALSE`. 

### Question 1: R Assignment Operators
Which of the following statements correctly describes the primary ways to assign variables in R?

```{r q1, echo=FALSE}
opts1 <- c(
  "Variables can only be assigned using the `=` operator.",
  "Variables can only be assigned using the `<-` operator.",
  answer = "Variables can be assigned using either the `<-` or the `=` operator.",
  "Variables are assigned using `->` for left-to-right assignment only."
)
```

`r longmcq(opts1)`

`r hide("Show Explanation")`
There are two primary ways to assign variables in R, using `<-` or `=`. While `->` exists for right-to-left assignment, `<-` and `=` are the standard primary assignment methods.
`r unhide()`

---

### Question 2: Comments in R
How do you write a single-line comment in an R script?
```{r q2, echo=FALSE}
opts2 <- c(
  "`// This is a comment`",
  "`/* This is a comment */`",
  answer = "`# This is a comment`",
  "`-- This is a comment`"
)
```

`r longmcq(opts2)`

### Question 3: Comments in R
If you create a variable named `myvar <- 10` and then call `MyVar`, what will happen?

```{r q3, echo=FALSE}
opts3 <- c(
  "It will return 10 because R is case-insensitive.",
  answer = "It will throw an 'object not found' error because R is case-sensitive.",
  "It will prompt you to correct the spelling.",
  "R will automatically convert `MyVar` to lowercase `myvar`."
)
```

`r longmcq(opts3)`


### Question 4: Understanding R Objects
What is the abstract term for anything that can be assigned to a variable in R, and which of these is not listed as a general object type in the primary enumeration?

```{r q4, echo=FALSE}
opts4 <- c(
  "An expression; functions are not listed.",
  "A function; integers are not listed.",
  answer = "An object; matrices are not listed as a general type under core object types.",
  "A variable; vectors are not listed."
)
```

`r longmcq(opts4)`

`r hide("Show Explanation")`
"Object" is the abstract term for anything assigned to a variable. Core enumerated object types include integers, vectors, and lists. While matrices are important data structures, they are built on vectors rather than listed as a standalone core type initially.
`r unhide()`

---

### Question 5: Data Types in R
Which of the following combinations correctly lists four main types of data in R and their corresponding examples?

```{r q5, echo=FALSE}
opts5 <- c(
  "Numerical (1000L), Character (TRUE), Integer (\"Table\"), Boolean (1)",
  answer = "Character (\"Table\"), Integer (1000L), Boolean (TRUE), Numerical (1)",
  "Numerical (\"Table\"), Character (1), Integer (TRUE), Boolean (1000L)",
  "Integer (1), Boolean (\"Table\"), Numerical (TRUE), Character (1000L)"
)
```

`r longmcq(opts5)`

`r hide("Show Explanation")`
In R:
* `"Table"` is a **Character** string.
* `1000L` (with the `L` suffix) is an **Integer**.
* `TRUE` is a **Boolean / Logical**.
* `1` is **Numeric**.
`r unhide()`

---

### Question 6: Characteristics of Vectors
Which statement accurately describes a crucial characteristic of vectors in R, particularly concerning arithmetic operations?

```{r q6, echo=FALSE}
opts6 <- c(
  "Vectors must always contain data of mixed types.",
  answer = "For calculations with vectors to be straightforward, they should generally be of the same length.",
  "Character vectors use acute accent keys for text strings.",
  "Data vectors are exclusively used for integers."
)
```

`r longmcq(opts6)`

`r hide("Show Explanation")`
Vector calculations operate element-by-element, so vectors should generally be of equal length to avoid unexpected recycling behavior.
`r unhide()`

---

### Question 7: Functions and Arguments
In an R function call like `log(x)`, what is the difference between `x` when it's provided in the call versus its definition within the function's structure?

```{r q7, echo=FALSE}
opts7 <- c(
  "x in the call is a 'formal argument', while in the function definition it's an 'actual argument'.",
  "Both are always 'formal arguments'.",
  answer = "x as provided in log(x) is an 'actual argument', connected to a 'formal argument' defined within the function.",
  "x is neither an actual nor formal argument; it's always an 'object'."
)
```

`r longmcq(opts7)`

`r hide("Show Explanation")`
* **Actual arguments** are the specific values/variables passed during a function call.
* **Formal arguments** are the parameters declared in the function's definition.
`r unhide()`

---

### Question 8: The Purpose of Factors
Why are 'factors' considered useful in R, particularly in statistical data analysis?

```{r q8, echo=FALSE}
opts8 <- c(
  "They allow vectors to contain mixed data types.",
  "They are primarily used for mathematical matrix operations.",
  answer = "They represent categorical variables, which is useful for indicating subdivisions of data and have a set of levels.",
  "They are a more efficient way to store numerical sequences."
)
```

`r longmcq(opts8)`

`r hide("Show Explanation")`
Factors store categorical variables (like experimental groups or levels of treatment) and track predefined levels.
`r unhide()`

---

### Question 9: Differentiating Lists and Vectors
What is the fundamental difference between a 'list' and a 'vector' in R regarding the types of objects they can contain?

```{r q9, echo=FALSE}
opts9 <- c(
  "Vectors can hold objects of different types, while lists can only hold objects of the same type.",
  answer = "Lists can hold collections of objects of different types, whereas a vector must contain elements of the same type.",
  "Both lists and vectors can hold any type of object without restriction.",
  "Lists are only for numeric data, and vectors are for character data."
)
```

`r longmcq(opts9)`

`r hide("Show Explanation")`
Atomic vectors require all elements to be of the exact same data type (homogenous). Lists are generic vectors that can store elements of varying types and shapes (heterogeneous).
`r unhide()`

---

### Question 10: Data Frames Structure
What is a 'data frame' in R, and what is its crucial characteristic regarding the length of its constituent components?

```{r q10, echo=FALSE}
opts10 <- c(
  "A data frame is a single vector of mixed data types; its components can have varying lengths.",
  "A data frame is a matrix that can only store numeric values.",
  answer = "A data frame is a list of vectors and/or factors of the same length, where data in the same position come from the same experimental unit.",
  "A data frame is an arbitrary collection of R objects without any length constraints."
)
```

`r longmcq(opts10)`

`r hide("Show Explanation")`
A data frame is structured like a table where each column is a vector or factor. All columns must have identical lengths (rows) for the data frame to be valid.
`r unhide()`

---

### Question 11: Printing Output: `print()` vs. `cat()`
When displaying a character vector in R, what is the primary difference between using `print()` and `cat()`?

```{r q11, echo=FALSE}
opts11 <- c(
  "print() displays only the first element, while cat() displays all elements.",
  "cat() always adds a newline character after each element, while print() does not.",
  answer = "print() includes quotation marks around character strings, whereas cat() displays them without quotation marks.",
  "print() is used for numeric data, and cat() is for character data."
)
```

`r longmcq(opts11)`

`r hide("Show Explanation")`
`print()` outputs formal string representations including quotes (e.g., `"John"`), whereas `cat()` concatenates and prints raw unquoted text (e.g., `John`).
`r unhide()`

---

### Question 12: Handling Missing Values (`NA`)
If you perform an arithmetic operation in R like `1 + NA`, what will be the typical result, and why?

```{r q12, echo=FALSE}
opts12 <- c(
  "The result will be 1, because NA is treated as 0 in arithmetic operations.",
  "The operation will produce an error, as NA cannot be used in arithmetic.",
  answer = "The result will be NA, because R treats NA as missing information and does not assume it is 0.",
  "The result will be NaN (Not a Number), indicating an undefined operation."
)
```

`r longmcq(opts12)`

`r hide("Show Explanation")`
R propagates `NA` values throughout calculations because operating on unknown data yields an unknown (`NA`) result.
`r unhide()`

---

## Part 2: Application Questions

### Question 13: Vector Creation and Naming
Which of the following R code snippets correctly creates vectors for temperatures and days, assigns the days as names to the temperatures, and prints the result?

```{r q13, echo=FALSE}
opts13 <- c(
  "A)<br><pre><code>temperature_readings = c(30.0, 28.0, 27.0, 28.5, 30.5, 29.0, 29.5)\ndays_of_week = c(\"Monday\", \"Tuesday\", \"Wednesday\", \"Thursday\", \"Friday\", \"Saturday\", \"Sunday\")\nlabels(temperature_readings) = days_of_week\ncat(temperature_readings)</code></pre>",
  answer = "B)<br><pre><code>temperature_readings <- c(30.0, 28.0, 27.0, 28.5, 30.5, 29.0, 29.5)\ndays_of_week <- c(\"Monday\", \"Tuesday\", \"Wednesday\", \"Thursday\", \"Friday\", \"Saturday\", \"Sunday\")\nnames(temperature_readings) <- days_of_week\nprint(temperature_readings)</code></pre>",
  "C)<br><pre><code>temp_readings <- list(30.0, 28.0, 27.0, 28.5, 30.5, 29.0, 29.5)\ndays <- c(\"Mon\", \"Tue\", \"Wed\", \"Thu\", \"Fri\", \"Sat\", \"Sun\")\nnames(temp_readings) <- days\nprint(temp_readings)</code></pre>",
  "D)<br><pre><code>temperature_readings <- c(30.0, 28.0, 27.0, 28.5, 30.5, 29.0, 29.5)\ndays_of_week <- c(\"Monday\", \"Tuesday\", \"Wednesday\", \"Thursday\", \"Friday\", \"Saturday\", \"Sunday\")\ntemperature_readings$names <- days_of_week\nprint(temperature_readings)</code></pre>"
)
```

`r longmcq(opts13)`

---

### Question 14: Vector Indexing
Given `temperature_readings <- c("Monday"=30.0, "Tuesday"=28.0, "Wednesday"=27.0, "Thursday"=28.5, "Friday"=30.5, "Saturday"=29.0, "Sunday"=29.5)`:

How would you extract the temperature for `"Wednesday"` using its name, and also remove the temperature readings for `"Saturday"` and `"Sunday"` from the vector?

```{r q14, echo=FALSE}
opts14 <- c(
  "temperature_readings[\"Wednesday\"] and temperature_readings[-c(6, 7)]",
  "temperature_readings[] and temperature_readings[-\"Saturday\", -\"Sunday\"]",
  answer = "temperature_readings[\"Wednesday\"] and temperature_readings[c(-6, -7)]",
  "temperature_readings$Wednesday and temperature_readings[1:5]"
)

cat(longmcq(opts14))
```

---

### Question 15: Sequence and Replicate Functions
Which R code snippet correctly creates a sequence from `0.5` to `10.0` with a step size of `0.75`, and repeats `my_data <- c("A", 10, TRUE)` so `"A"` appears once, `10` twice, and `TRUE` three times?

```{r q15, echo=FALSE, results='asis'}
opts15 <- c(
  "seq(0.5, 10.0, by = 0.75)\nmy_data <- c(\"A\", 10, TRUE)\nrep(my_data, each = c(1, 2, 3))",
  "seq(from = 0.5, to = 10.0, by = 0.75)\nmy_data <- c(\"A\", 10, TRUE)\nrep(my_data, times = c(1, 2, 3))",
  "sequence(0.5, 10.0, 0.75)\nmy_data <- c(\"A\", 10, TRUE)\nrep(my_data, c(1, 2, 3))",
  answer = "seq(from = 0.5, to = 10.0, by = 0.75)\nmy_data <- c(\"A\", 10, TRUE)\nrep(my_data, c(1, 2, 3))"
)
```

`r longmcq(opts15)`
---

### Question 16: Matrix Creation and Transposition
Which R code snippet correctly creates a matrix with 4 rows and 3 columns filled row-wise with numbers 1 to 12, sets row names, and transposes it?

```{r q16, echo=FALSE, results='asis'}
opts16 <- c(
  "my_matrix <- matrix(1:12, ncol = 3, byrow = TRUE)\nrownames(my_matrix) <- paste0(\"Row\", 1:4)\nt(my_matrix)",
  "my_matrix <- matrix(1:12, nrow = 4, byrow = FALSE)\nrow.names(my_matrix) <- c(\"Row1\", \"Row2\", \"Row3\", \"Row4\")\ntranspose(my_matrix)",
  answer = "my_matrix <- matrix(1:12, nrow = 4, ncol = 3, byrow = TRUE)\nrownames(my_matrix) <- c(\"Row1\", \"Row2\", \"Row3\", \"Row4\")\nt(my_matrix)",
  "my_matrix <- as.matrix(1:12, rows = 4, cols = 3, fill.by.row = TRUE)\nrownames(my_matrix) <- LETTERS[1:4]\nmy_matrix %T%"
)
```

`r longmcq(opts16)`

---

### Question 17: Combining Vectors into Matrices
Which snippet correctly combines `math_scores`, `science_scores`, and `art_scores` into columns with named headings `"Math"`, `"Science"`, and `"Art"`?

```{r q17, echo=FALSE, results='asis'}
opts17 <- c(
  "student_scores_by_col <- rbind(math_scores, science_scores, art_scores)\ncolnames(student_scores_by_col) <- c(\"Math\", \"Science\", \"Art\")",
  "student_scores_by_col <- matrix(c(math_scores, science_scores, art_scores), ncol = 3, byrow = FALSE)\ncolnames(student_scores_by_col) <- c(\"Math\", \"Science\", \"Art\")",
  answer = "student_scores_by_col <- cbind(\"Math\" = math_scores, \"Science\" = science_scores, \"Art\" = art_scores)",
  "student_scores_by_col <- list(Math = math_scores, Science = science_scores, Art = art_scores)"
)
```

`r longmcq(opts17)`

---

### Question 18: Factor Application
Given `rating_values <- c(1, 0, 2, 1, 3, 0)`, you create a factor `customer_rating` from it with levels `"Poor"`, `"Fair"`, `"Good"`, `"Excellent"` representing 0, 1, 2, and 3 respectively. What would `as.numeric(customer_rating)` return?

```{r q18, echo=FALSE, results='asis'}
opts18 <- c(
  "c(\"Fair\", \"Poor\", \"Good\", \"Fair\", \"Excellent\", \"Poor\")",
  "c(0, 1, 2, 1, 3, 0)",
  "c(\"Poor\", \"Fair\", \"Good\", \"Excellent\")",
  answer = "c(2, 1, 3, 2, 4, 1)"
)
```

`r longmcq(opts18)`

`r hide("Show Explanation")`
Factors store levels internally as integers starting at 1. Since `"Poor"` (0) is the first level, it gets coded as `1`, `"Fair"` (1) becomes `2`, `"Good"` (2) becomes `3`, and `"Excellent"` (3) becomes `4`.
`r unhide()`

---

### Question 19: Data Frame Creation and Access
Which code snippet creates a data frame `weather_df`, prints the `days_of_week` column using two valid methods, and prints the first row?

```{r q19, echo=FALSE, results='asis'}
opts19 <- c(
  "weather_df <- data.frame(temperature_readings, days_of_week)\nprint(weather_df$days_of_week)\nprint(weather_df[, 2])\nprint(weather_df[1])",
  answer = "weather_df <- data.frame(temperature_readings, days_of_week)\nprint(weather_df$days_of_week)\nprint(weather_df[, \"days_of_week\"])\nprint(weather_df[1, ])",
  "weather_df <- list(temp = temperature_readings, days = days_of_week)\nprint(weather_df[[\"days\"]])\nprint(weather_df[[2]])\nprint(weather_df[1])",
  "weather_df <- data.frame(temp = temperature_readings, day = days_of_week)\nprint(weather_df[2])\nprint(weather_df[c(T,T,F,F,F,F,F), ])\nprint(weather_df[1])"
)
```

`r longmcq(opts19)`

---

### Question 20: Conditional Selection for Vectors
Which code snippet extracts temperatures greater than `29.0`, and then extracts temperatures that are `<= 28.0` OR `> 30.0`?

```{r q20, echo=FALSE, results='asis'}
opts20 <- c(
  "temperature_readings[temperature_readings >= 29.0]\ntemperature_readings[temperature_readings <= 28.0 & temperature_readings > 30.0]",
  answer = "temperature_readings[temperature_readings > 29.0]\ntemperature_readings[temperature_readings <= 28.0 | temperature_readings > 30.0]",
  "temperature_readings[which(temperature_readings > 29.0)]\ntemperature_readings[any(temperature_readings <= 28.0, temperature_readings > 30.0)]"
)
```

`r longmcq(opts20)`
---

### Question 121: Conditional Selection for Data Frames
Using `weather_df`, how do you select all rows where `temperature_readings > 29.5`, and rows where `rainfall_mm > 5.0`?

```{r q21, echo=FALSE, results='asis'}
opts21 <- c(
  "weather_df[temperature_readings > 29.5, ]\nweather_df[rainfall_mm > 5.0, ]",
  answer = "weather_df[weather_df$temperature_readings > 29.5, ]\nweather_df[weather_df$rainfall_mm > 5.0, ]",
  "subset(weather_df, temperature_readings > 29.5)\nsubset(weather_df, rainfall_mm > 5.0)",
  "weather_df[weather_df[[\"temperature_readings\"]] > 29.5]\nweather_df[weather_df[[\"rainfall_mm\"]] > 5.0]"
)
```

`r longmcq(opts21)`
---

### Question 22: Handling Missing Values
You have `sales_data <- c(100, 150, NA, 120, 200, NA)`. Which code correctly checks for missing values and calculates the sum excluding `NA`s?

```{r q22, echo=FALSE, results='asis'}
opts22 <- c(
  "a)<br><code>any(is.na(sales_data))</code><br><code>sum(sales_data, na.omit = TRUE)</code>",
  "b)<br><code>is.na(sales_data)</code><br><code>sum(sales_data, ignore.na = TRUE)</code>",
  answer = "c)<br><code>is.na(sales_data)</code><br><code>sum(sales_data, na.rm = TRUE)</code>",
  "d)<br><code>sales_data == NA</code><br><code>sum(sales_data[complete.cases(sales_data)])</code>"
)
```

`r longmcq(opts22)`

`r hide("Show Explanation")`
`is.na()` tests each element for missingness, and `na.rm = TRUE` tells `sum()` to ignore `NA` values during addition.
`r unhide()`



