---
title: "Introduction to R and RStudio"
output: webexercises::webexercises_default
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(webexercises)
```


```{r, echo = FALSE, results='asis'}
# Uncomment to change widget colours:
#style_widgets(incorrect = "goldenrod", correct = "purple")
```


## Longer MCQs (`longmcq()`)

When your answers are very long, sometimes a drop-down select box gets formatted oddly. You can use `longmcq()` to deal with this. Since the answers are long, It's probably best to set up the options inside an R chunk with `echo=FALSE`. 

### Question 1: R Assignment Operators
Which of the following statements correctly describes the primary ways to assign variables in R?

```{r q1, echo=FALSE}
opts1 <- c(
  "Variables can only be assigned using the `=` operator.",
  "Variables can only be assigned using the `<-` operator.",
  answer = "Variables can be assigned using either the `<-` or the `=` operator.",
  "Variables are assigned using `->` for left-to-right assignment only."
)
```

`r longmcq(opts1)`

`r hide("Show Explanation")`
There are two primary ways to assign variables in R, using `<-` or `=`. While `->` exists for right-to-left assignment, `<-` and `=` are the standard primary assignment methods.
`r unhide()`

